\section{September 2, 2021} 
What is the point of abstraction? Sets allow us to not worry about the elements themselves (like apples or oranges), but about the structure of the set itself. Category theory is the next level of abstraction, where the objects have relations that keep track of structure.

There may be some constructions like exterior algebras, tensor products, etc that we wish to construct on finite dimensional vector spaces. Categorically, we can use the correspondence with the ``skeleton'' category $\mathsf{Vect} _{\mathrm{fd}}^{\R}\xleftrightarrow{\sim} \mathsf{Euc} $, forming a \textbf{full subcategory}- a subcategory that loses some objects but remembers \emph{all} morphism.

\subsection{Initial and terminal objects}
Recall the poset category ($P, \leq$). We can ask for a \emph{maximal} element $M \in P$ such that for all $x \in P$, $x \leq M$, or a \emph{minimal} element $m$, where $m\leq x$ for all $x \in P$.
\begin{definition}[]
    Let $\mathcal{C} $ be a category. We say $x \in \mathcal{C} $ is a \textbf{final object} (or \emph{terminal object}) of $\mathcal{C} $ if for all $y \in C$, there exists a unique map $y \to x$. Similarly, an \textbf{initial object} $m \in \mathcal{C} $ is an object such that for all $y \in \mathcal{C} $, there exists a unique $m \to y$.
\end{definition}
So in a poset, maximal elements are final and minimal objects are initial (if they exist). In general, universal properties are not guaranteed to exist: but if they do, they are unique.

\begin{prop}
    If a final object exists, it is unique up to unique isomorphism. In other words, if $M_1,M_2$ are both final objects of $\mathcal{C} $, there is a unique $M_1 \xrightarrow{\sim} M_2$.
\end{prop}
\begin{proof}
    Assume $M_1$ is final, so there exists a unique map $M_2 \to M_1$. Similarly, there exists a unique map $M_1 \to M_2$. Composing these two maps gives a unique map $M_1 \to M_1$, which must be the identity.
\end{proof}
\begin{example}
    Here are some examples of initial and final objects:
    \begin{itemize}
    \setlength\itemsep{-.2em}
\item In $\mathsf{Set} $, the singleton $\{*\} $ is final while the empty set $\O$ is initial.
\item In $\mathsf{Vect} _{\R}$ or $\mathsf{Ab} $, the zero vector space/trivial group is both initial (linear maps send identities to identities) and final. 

    This gives rise to the idea of a \textbf{zero object}, which is both initial and final. ``Geometric'' categories like $\mathsf{Top} $ and $\mathsf{Set} $ tend not to have zero objects, but ``linear'' categories like the ones above do.
\item In $\mathsf{Ring} $, 0 is final while $\Z$ is initial (it has zero and one).
\item The category of fields does not have an initial or terminal object.
    \end{itemize}
\end{example}

\subsection{Localization}

We start with $R$ a ring, and we choose a multiplicative subset $S \subseteq R$ (containing 1, closed under $\cdot $). We want to add inverse to $s \in S$ to the ring $R$. Consider the ``undercategory'' $\mathsf{Ring} $ {\color{red}todo:see other notes??} with objects rings $T$ with a morphism $R \xrightarrow{\varphi _T} T$, with morphisms {\color{red}todo:??} 
\begin{definition}[]
    The \textbf{localization} $S ^{-1}R$ (if it exists) is the initial object of $\mathcal{C} _{R,S}$. In other words, {\color{red}todo:?} 
\end{definition}
The way of constructing localizations by hand is by some equivalence relations $\frac{r}{s}\sim \frac{r'}{s'}$, which leads to $rs'-r's=0$. However, there may be zero divisors, so instead mod out by $(rs'-r's)s''=0$, or $rs'-r's=\frac{0}{s''}$.

\begin{note}
    Think of rings as function rings, with a space associated (eg smooth functions on $\R^n $).
\end{note}

\subsection{Completion}
Let $R$ be a ring (eg $\Z$) and $I \subseteq R$ an ideal $(p)$. {\color{red}todo:?? go back and finish}  Consider $R /I^2 \to R/I$. Completion is the limit on the left. Consider $\mathcal{C} $, the category of all rings $T$ with compatible maps $T \to  R / I ^n $ for rall $n$.

\begin{definition}[]
    The completion $\hat{R}_I$ is the final object of $\mathcal{C} $ if it exists.
\end{definition}
