\section{The Fundamental Group}
OK guys, let's decompose big spaces into smaller ones and compute their fundamental groups. These notes follow Hatcher \S 1.2, Lee \S 10, and May \S 2.7.
\subsection{Fundamental group of the circle(todo)}
If this is a first introduction to fundamental groups, then our first fundamental group of real interest is $\pi_1(S^{1} )=\Z$. Before we do this, let's do a quick calculation to show $\pi_1(\R)=0$. Take the origin as a convenient basepoint. Define $k \colon \R\times I \to \R$ by $k(s,t)=(1-t)s$. Then $k$ is a homotopy from the identity to the constant map at $0$. For a loop $f \colon I \to \R$ at $0$, define $h(s,t)=k(f(s),t).$ Then $f$ is equivalent to a constant $c_0$ by the homotopy $h$.
\orbreak
Now let's talk about circles: we can view $S^{1} $ as the circle group (let's denote it $U^1$), that is, \[
U^1 = \{z\in \C\mid |z|=1\}.
\] Multiplication is continuous, so this is a topological group. Take the identity $1$ as a convenient basepoint for  $S^{1} $.
\begin{theorem}
    We have the fundamental group of a circle isomorphic to the integers, that is, \[
        \pi_1(S^{1} ,1)\cong\Z.
    \] 
\end{theorem}
\begin{proof}
    For all $n\in \Z$, define a loop $f_n$ in $S^{1} $ by $f_n(s)=e^{2\pi i ns}$. This is the same as composing the map $I\to S^{1},\, s\mapsto e^{2\pi is}$ and the $n$th power map on $S^{1} $. If we identiy the boundary points $0$ and $1$ of $I$, then the first map $(I \to S^{1} )$ induces the evident identification of $I / \partial I$ with $S^{1} $. By complex exponentiation, we have $[f_m][f_n]=[f_{m+n}]$: define a homomorphism $i \colon \Z \to \pi_1(S^{1} ,1)$ by $i(n)=[f_n]$. We claim $i$ is an isomorphism. The main idea is to use the fact that (locally) $S^{1} $ looks like $\R$, ie, $S^{1} $ is a $1$-manifold.

    ***FINISH LATER***
\end{proof}


\subsection{The van Kampen Theorem (Hatcher)}
Let's take a space $X$ and say it's the union of path-connected open subsets $A_{\alpha}$, each of which contains the basepoint $x_0\in X$. Then the homomorphisms $j_{\alpha} \colon \pi_1(A_{\alpha}) \to \pi_1(X)$ induced by the inclusions $A_{\alpha}\hookrightarrow X$ extend to a homomorphism $\Phi \colon *_{\alpha}\pi_1(A_{\alpha}) \to \pi_1(X)$. The van Kampen theorem will say that $\Phi$ is often onto but in general, we can expect $\Phi$ to have a nontrivial kernel. 

For if $i_{\alpha\beta} \colon  \pi_1(A_{\alpha}\cap A_{\beta}) \to \pi_1(A_{\alpha})$ is the homomorphism induced by the inclusion $A_{\alpha}\cap A_{\beta}\hookrightarrow A_{\alpha}$ then $j_{\alpha}i_{\alpha\beta}=j_{\beta}i_{\beta\alpha}$, both of these compositions being induced by the inclusion $A_{\alpha}\cap A_{\beta}\hookrightarrow X$, so the kernel of $\Phi$ contains all the elements of the form $i_{\alpha\beta}(\omega)i_{\beta\alpha}(\omega)^{-1}$ for $\omega \in \pi_1(A_{\alpha}\cap A_{\beta})$. 

Van Kampen says under fairly broad hypotheses that this determines all of $\Phi$.

\begin{theorem}
    If $X$ is the union of path-connected open sets $A_{\alpha}$ each containing the basepoint $x_0\in X$ and if each intersection $A_{\alpha}\cap A_{\beta}$ is path-connected, then the homomorphism \[
        \Phi \colon *_{\alpha}(A_{\alpha}) \to \pi_1(X)
    \] is onto. Furthermore, if each intersection $A_{\alpha}\cap A_{\beta}\cap A_{\gamma}$ is path-connected, then the kernel of $\Phi$ is the normal subgroup $N$ generated by all elements of the form $i_{\alpha\beta}(\omega)i_{\beta\alpha}(\omega)^{-1}$ for $\omega \in \pi_1(A_{\alpha}\cap A_{\beta})$, and hence $\Phi$ induces an isomorphism \[
    \pi_1(X)=*_{\alpha}\pi_1(A_{\alpha}) /N.
    \] 
\end{theorem}
\begin{example}[Wedge Sums]
    I like the visual of the wedge sum but the terminology of the smash product. If only we could keep the \texttt{\textbackslash vee} symbol ($\vee$) and say we ``smash the spaces together'' at a point. 

    We define the wedge sum $\bigvee_{\alpha}X_{\alpha}$ with basepoints $x_{\alpha}\in X_{\alpha}$ as the disjoint union $\amalg_{\alpha}X_{\alpha}$ with all the basepoints $x_{\alpha}$ identified to a single point. If each $x_{\alpha}$is a deformation retract of an open neighborhood $U_{\alpha }$ in $X_{\alpha }$, then $X_{\alpha }$ is a deformation retract of its open neighborhood $A_{\alpha }=X_{\alpha }\bigvee_{\beta\neq\alpha }U_{\beta}$. The intersection of two or more distinct $A_{\alpha }$'s is $\bigvee_{\alpha }U_{\alpha }$, which deformation retracts to a point. Then by van Kampens theorem, \[
        \Phi \colon *_{\alpha }\pi_1(X_{\alpha }) \to \pi_1(\bigvee_{\alpha }X_{\alpha })
    \] is an isomorphism, provided each $X_{\alpha }$ is path-connected, hence also each $A_{\alpha }$. Therefore for a wedge sum of circles, $\pi_1(\bigvee_{\alpha }S_{\alpha }^{1})$ is a free group, the free product of copies of $\Z$.
\end{example}
\orbreak
I know it always helps to see something done somewhere else. For me, the above definition fails to make any sense at all whatsoever. So, let's revisit van Kampens from two more lens: one from the words of Lee (\emph{Introduction to Topological Manifolds}) and another from the categorical perspective.
\subsection{The van Kampen Theorem (Lee)}
Let's say we have a space $X$ that's made up of the union of two open subsets $U,V\subseteq X$. We know how to compute the fundamental groups of $U,V$, and $U\cap V$ (each of which is path-connected). Every loop can be written as a product of loops in $U$ or $V$ (visualized as the free product $\pi_1(U)*\pi_1(V)$), but any loop in $U\cap V$ only represents a single element of $\pi_1(X)$, even though it represents two distinct elements of the free product (one in $\pi_1(U)$ and one in $\pi_1(V)$). So $\pi_1(X)$ can be though of as the quotient of this free product modulo some relations from $\pi_1(U\cap V)$ that express this redundancy.

Let's do some setup so we can state a precise version of van Kampens. Let $X$ be a topological space and $U,V\subseteq X$ such that $U \cup V=X$ and $U\cap V\neq \O$. Let $q\in U\cap V$. Then the four inclusion maps shown below,
            \begin{figure}[H]
                \centering
                \begin{tikzcd}
                                         & U \arrow[rd, "k"]  &   \\
U\cap V \arrow[ru, "i"] \arrow[rd, "j"'] &                    & X \\
                                         & V \arrow[ru, "l"'] &  
\end{tikzcd}
            \end{figure}
            induce fundamental group homomorphisms as shown below.
                        \begin{figure}[H]
                \centering
                \begin{tikzcd}
                                                        & {\pi_1(U,q)} \arrow[rd, "k_*"] &              \\
{\pi_1(U\cap V,q)} \arrow[ru, "i_*"] \arrow[rd, "j_*"'] &                                & {\pi_1(X,q)} \\
                                                        & {\pi1(V,q)} \arrow[ru, "l_*"'] &             
\end{tikzcd}
            \end{figure}
            Now insert the free product $\pi_1(U,q)*\pi_1(V,q)$ in the middle of the diagram and let $\iota_U \colon \pi_1(U,q) \hookrightarrow \pi_1(U,q)*\pi_1(V,q)$ and $\iota_V \colon \pi_1(V,q) \hookrightarrow \pi_1(U,q)*\pi_1(V,q)$ be the canonical injections. By the characteristic property (unique induced homomorphisms) of the free product, $k_*$ and $l_*$ induce a homomorphism $\Phi \colon \pi_1(U,q)*\pi_1(V,q) \to \pi_1(X,q)$ such that the right half of the following diagram commutes:

                        \begin{figure}[H]
                \centering
                \begin{tikzcd}
                                                                                & {\pi_1(U,q)} \arrow[rd, "k_*"] \arrow[d, "\iota_U" description]  &              \\
{\pi_1(U\cap V, q)} \arrow[ru, "i_*"] \arrow[rd, "j_*"'] \arrow[r, "F", dotted] & {\pi_1(U,q)*\pi_1(V,q)} \arrow[r, "\Phi"]                        & {\pi_1(X,q)} \\
                                                                                & {\pi_1(V,q)} \arrow[ru, "l_*"'] \arrow[u, "\iota_V" description] &             
\end{tikzcd}
            \end{figure}
            Finally, define a map $F \colon \pi_1(U\cap V,q) \to \pi_1(U,q)*\pi_1(V,q)$ by setting $F(\gamma)=(i_*\gamma)^{-1}(j_*\gamma)$\footnote{$F$ is not a homomorphism.}. Let $\overline{F(\pi_1(U\cap V,q))}$ denote the \emph{normal closure}\footnote{the \emph{normal closure} of a set means the smallest normal subgroup that contains such set.} of the image of $F$ in $\pi_1(U,q)*\pi_1(V,q)$.

\begin{theorem}[Seifert-Van Kampen]
    Let $X$ be a topological space. Suppose $U,V\subseteq X$ are open, $U\cap V=X$, and $U,V$, and $U\cap V$ are path-connected. Then for any $q\in U\cap V$, the homomorphism $\Phi$ is surjective, and its kernel is $\overline{F(\pi_1(U\cap V,q))}$. Therefore we have \[
        \pi_1(X,q)\cong \pi_1(U,q)*\pi_1(V,q) \big/ \overline{F(\pi_1(U\cap V,q))}.
    \] 
\end{theorem}
When the fundamental groups in question are finitely presented, the theorem has a useful reformulation in terms of generators and relations.
\begin{cor}
    In addition to the hypothesis of van Kampen, assume that the fundamental groups of $U,V$, and $U\cap V$ have the following finite presentations:
    \begin{gather*}
\pi_1(U,q)\cong \langle \alpha_1,\cdots,\alpha_m \mid \rho_1,\cdots,\rho_r \rangle;\\
\pi_1(V,q)\cong\langle \beta_1,\cdots,\beta_n \mid \sigma_1,\cdots,\sigma_s \rangle; \\
\pi_1(U\cap V,q)\cong \langle \gamma_1,\cdots,\gamma_p \mid \tau_1,\cdots,\tau_t \rangle .
    \end{gather*}
    Then $\pi_1(X,q)$ has the presentation \[
        \pi_1(X,q)\cong \langle \alpha_1,\cdots,\alpha_m,\,\beta_1,\cdots,\beta_n \mid \rho_1,\cdots,\rho_r,\,\sigma_1,\cdots,\sigma_s,\,u_1=v_1,\cdots,u_p=v_p \rangle 
    \] where for each $a=1,\cdots,p$, $u_a$ is an expression for $i_*\gamma_a\in \pi_1(U,q)$ in terms of the generators $\{\alpha_1,\cdots,\alpha_m\} $, and $v_a$ similarly expresses $j_*\gamma_a\in \pi_1(V,q)$ in terms of $\{\beta_1,\cdots,\beta_n\} $.
\end{cor}
\subsection{The fundamental groupoid}
We backtrack a little to talk about categorical nonsense. This doesn't fit too well with the section on category theory, so it's here. These will follow May \S 2.5.
\orbreak
We often talk of pointed spaces, but it would to nice to talk about spaces without making such a choice. We define the fundamental groupoid $\prod_{}^{} (X) $ of a space $X$ to be the category whose objects are the points of $X$ and whose morphism $x \to y$ are the equivalence classes of paths from $x$ to $y$\footnote{Recall Example 1.4 of a group being realized as a category with all its arrows isomorphisms.}. Then the set of endomorphisms of the object $x$ is the fundamental group $\pi_1(X,x)$.

We say ``groupoid'' because a group is simply a groupoid with only one object (the class of morphisms or symmetries on an object). However, the category of groupoids has several objects. We also defined groupoids as categories whose morphisms are all isomorphisms. If morphisms are functors, then we have the category $\mathsf{Grpd} $ of groupoids. So we can see $\prod $ as a functor $\mathsf{Top_*} \to \mathsf{Grpd} $.

Let's talk about skeletons. We have the skeleton of a category $\mathcal{C} $ denoted by $\operatorname{sk}(\mathcal{C}) $. This is a ``full'' subcategory with one object from each isomorphism class of objects of $\mathcal{C} $, ``full'' meaning that the morphisms between two objects of $\operatorname{sk}(\mathcal{C} )$ are all of the morphisms between these objects in $\mathcal{C} $. The inclusion functor $J \colon \operatorname{sk}(\mathcal{C} ) \to \mathcal{C} $ is an equivalence of categories. We can find an inverse functor $F \colon \mathcal{C}  \to \operatorname{sk}(\mathcal{C} )$ by letting $F(A)$ be the unique object in $\operatorname{sk}(\mathcal{C} )$ that is isomorphic to $A$, choosing an isomorphism $\alpha _A \colon A \to F(A)$, and defining $F(f)=\alpha _B \circ f \circ \alpha_A^{-1}$ for a morphism $f \colon  A \to B$ in $\mathcal{C} $. Choose $\alpha $ to be the identity morphisms if $A\in \operatorname{sk}(\mathcal{C} )$, then $FJ=\operatorname{id}_{\operatorname{sk}(\mathcal{C} )}$; the $\alpha _A$ specify a natural isomorphism $\alpha \colon \operatorname{id} \to JF $.

A category is connected if any two objects can be connected by a sequence of morphisms. Then a groupoid is connected iff any two of its objects are isomorphic. The group of endomorphisms of any object $C$ is then a skeleton of $\mathcal{C} $, so we can generalize our results about skeletons to give the relationship between a fundamental group and a fundamental groupoid of a path connected space $X$.
\begin{prop}
    Let $X$ be a path connected space. Then for each $x\in X$, the inclusion $\pi_1(X,x) \to \prod (X)$ is an equivalence of categories.
\end{prop}
\begin{proof}
    View $\pi_1(X,x)$ as a groupoid with on object $x$: then $\pi_1(X,x)$ is a skeleton of $\prod (X)$ and we are done.
\end{proof}
May's presentation and proofs are very concise and elegant. I like this.



