\section{Category Theory}
Today we talk about abstract nonsense! These notes will follow Evan Chen's Napkin \S 60 and May's ``A Concise Course in Algebraic Topology'' \S 2.

\subsection{Motivation}
Why do we talk about categories? Categories rise from objects (sets, groups, topologies) and maps between them (bijections, isomorphisms, homeomorphisms). Algebraic topology speaks of maps from topologies to groups, which makes maps between categories a suitable tool for us.
\begin{example}
    Here are some examples of morphisms between objects:
    \begin{itemize}
        \item A bijective homomorphism between two groups $G$ and $H$ is an isomorphism. What also works is two group homomorphisms $\phi \colon G \to H$ and $\psi \colon H \to G$ which are mutual inverses, that is $\phi \circ \psi = \operatorname{id}_H$ and $\psi \circ \phi=\operatorname{id}_G$.
        \item Metric (or topological) spaces $X$ and  $Y$ are isomorphic if there exists a continuous bijection $f \colon  X \to Y$ such that $f^{-1}$ is also continuous.
        \item Vector spaces $V$ and $W$ are isomorphic if there is a bijection $T \colon V \to W$ that's a linear map (aka, $T$ and $T^{-1}$ are linear maps).
        \item Rings $R$ and $S$ are isomorphic if there is a bijective ring homomorphism $\phi$ (or two mutually inverse ring homomorphism).
    \end{itemize}
\end{example}

\subsection{Categories}
\begin{definition}[Category]
    A \textbf{category} $\mathcal{A}$ consists of 
    \begin{itemize}
        \item A class of \textbf{objects} , denoted $\operatorname{obj}(\mathcal{A}).$ 
        \item For any two objects $A_1,A_2\in \operatorname{obj}(\mathcal{A})$, a class of \textbf{arrows} (also called  \textbf{morphisms} or \textbf{maps} between them). Let's denote the set of arrows by $\operatorname{Hom}_{\mathcal{A}}(A_1,A_2)$.
        \item For any $A_1,A_2,A_3\in \operatorname{obj}(\mathcal{A})$, if $f \colon A_1 \to A_2$ is an arrow and $g \colon A_2 \to A_3$ is an arrow, we can compose the two arrows to get $h=g\circ f \colon A_1 \to A_3$ an arrow, represented in the \textbf{commutative diagram} below: 
            \begin{figure}[H]
                \centering
    \begin{tikzcd}
A_1 \arrow[rd, "h", dotted] \arrow[r, "f"] & A_2 \arrow[d, "g"] \\
                                           & A_3               
\end{tikzcd}
            \end{figure}
        The composition operation can be denoted as a function \[
            \circ \colon \operatorname{Hom}_{\mathcal{A}}(A_2,A_3)\times \operatorname{Hom}_{\mathcal{A}}(A_1,A_2) \to \operatorname{Hom}_{\mathcal{A}}(A_1,A_3)
        \] for any three objects $A_1,A_2,A_3$. Composition must be associative, that is, $h\circ(g\circ f)=(h\circ g)\circ f.$ In the diagram above, we say $h$ \textbf{factors} through $A_2$.
    \item Every object $A\in \operatorname{obj}_{\mathcal{A}}$ has a special \textbf{identity arrow} $ \operatorname{id}_{\mathcal{A}}$. The identity arrow has the expected properties $\operatorname{id}_{\mathcal{A}}\circ f=f$ and $f\circ \operatorname{id}_{\mathcal{A}}=f$.
    \end{itemize}
\end{definition}
\begin{note}
    We can't use the word ``set'' to describe the class of objects because of some weird logic thing (there is no set of all sets). But you can think of a class as a set.
\end{note}
From now on, $A\in \mathcal{A}$ is the same as $A\in \operatorname{obj}(\mathcal{A})$. A category is \textbf{small} if it has a set of objects, and \textbf{locally small} if $\operatorname{Hom}_{\mathcal{A}}(A_1,A_2)$ is a set for any $A_1,A_2\in \mathcal{A}$.

\begin{example}[Basic Categories]
    Here are some basic examples of categories:
    \begin{itemize}
        \item We have the category of groups $\mathsf{Grp}.$ 
            \begin{itemize}
                \item The objects of $\mathsf{Grp}$ are groups.
                \item The arrows of $\mathsf{Grp}$ are group homomorphisms.
                \item The composition of $\mathsf{Grp}$ is function composition.
            \end{itemize}
        \item Describe the category $\mathsf{CRing}$ (of commutative rings) in a similar way.
        \item Consider the category $\mathsf{Top}$ of topological spaces, whose arrows are continuous maps between spaces.
        \item Also consider the category $\mathsf{Top}_{*}$ of topological spaces with a distinguished basepoint, that is, a pair $(X,x_0), \, x_0\in X$. Arrows are continuous maps $f \colon X \to Y$ with $f(x_0)=y_0$.
        \item Similarly, the category of (possibly infinite-dimensional) vector spaces over a field $k$ $\mathsf{Vect}_{k}$ has linear maps for arrows. There is even a category $\mathsf{FDVect}_{k}$ of finite-dimensional vector spaces.
        \item Finally, we have a category $\mathsf{Set}$ of sets, arrows denote any map between sets.
    \end{itemize}
\end{example}
\begin{definition}[Isomorphism]
    An arrow $A_1\overset{f}{\to }A_2$ is an \textbf{isomorphism} if there exists $A_2\overset{g}{\to }A_1$ such that $f\circ g=\operatorname{id}_{A_2}$ and $g\circ f=\operatorname{id}_{A_1}$. We say $A_1$ and $A_2$ are \textbf{isomorphic}, denoted $A_1\cong A_2$.
\end{definition}
\begin{remark}
    In the category $\mathsf{Set}$, $X\cong Y \iff |X|=|Y|$.
\end{remark}
In other fields, we can tell a lot about the structure of an object by looking at maps between them. In category theory, we \emph{only} look arrows, and ignore what the objects themselves are.
\begin{example}[Posets are Categories]
   Let $\mathcal{P} $ be a poset. Then we can construct a category $P$ for it as follows:
   \begin{itemize}
       \item The objects of $P$ are elements of $\mathcal{P} $.
    \item We define the arrows of $P$ as follows:
        \begin{itemize}
            \item For every object $p\in P$, we add an identity arrow $\operatorname{id}_p$, and 
            \item For any pair of distinct objects $p\leq q$, we add a single arrow $p\to q$. 
        \end{itemize}
        There are no other arrows.
    \item We compose arrows in the only way possible, examining the order of the first and last object.
   \end{itemize}
   Here's a figure depicting the category of a poset $\mathcal{P} $ on four objects $\{a,b,c,d\} $ with $a\leq b$ and $a\leq c\leq d$.
            \begin{figure}[H]
                \centering
                \begin{tikzcd}
                                                                          & {[a]} \arrow[dd, "a\leq d" description] \arrow[rd, "a\leq c" description] \arrow[ld, "a\leq b" description] \arrow["\operatorname{id}_a"', loop, distance=2em, in=125, out=55] &                                                                                                            \\
{[b]} \arrow["\operatorname{id}_b"', loop, distance=2em, in=215, out=145] &                                                                                                                                                                                & {[c]} \arrow[ld, "c\leq d" description] \arrow["\operatorname{id}_c"', loop, distance=2em, in=35, out=325] \\
                                                                          & {[d]} \arrow["\operatorname{id}_d"', loop, distance=2em, in=305, out=235]                                                                                                      &                                                                                                           
\end{tikzcd}
            \end{figure}
            Note that no two distinct objects of a poset are isomorphic.
\end{example}
This shows that categories don't have to refer to just structure preserving maps between sets (these are called ``concrete categories''.
\begin{example}[Groups as a category with one object]
   A group $G $ can be thought of as a category $\mathcal{G} $ with one object $*$, all of whos arrows are isomorphisms. 

            If the universe were structured differently and kids learned category theory before groups, symmetries transforming $X$ into itself would be a natural extension of categories that transform $X$ into other objects, a special case in which all the maps are invertible. Alas, this is not the right timeline.
\end{example}
\begin{example}[Deriving Categories]
We can make categories from other categories!
   \begin{enumerate}
       \item[(a)] Given a category $\mathcal{A} $, we can construct the \textbf{opposite category} $A^{\text{op}}$, which is the same as $\mathcal{A} $ but with all the arrows reversed.
       \item[(b)] Given categories $\mathcal{A} $ and $\mathcal{B} $, we can construct the \textbf{product category}  $\mathcal{A} \times \mathcal{B} $ as follows: the objects are pairs $(A,B)$ for $A\in \mathcal{A} $, $B\in \mathcal{B} $, and the arrows from $(A_1,B_1)$ to $(A_2,B_2)$ are pairs \[
               \left( A_1 \overset{f}{\to }A_2, B_1\overset{g}{\to }B_2 \right) .
       \] The composition is just pairwise composition, and the identity is the pair of identity functions on $A$ and $B$.
   \end{enumerate} 
\end{example}

\subsection{Special objects in categories}
Some categories have things called \emph{initial objects}. For example the empty set $\O$, the trivial group, the empty space, initial element in a poset, etc. More interestingly: the initial object of $\mathsf{CRing} $ is the ring $\Z$.
\begin{definition}[Initial object]
    An \textbf{initial object} of $\mathcal{A} $ is an object $A_{\text{init}}\in \mathcal{A} $ such that for any $A\in \mathcal{A} $ (possibly $A=A_{\text{init}}$), there is exactly one arrow from $A_{\text{init}}$ to $A$.
\end{definition}
The yang to this yin is the \emph{terminal object}:
\begin{definition}[Terminal object]
    A \textbf{terminal object} of $\mathcal{A} $ is an object $A_{\text{final}}\in \mathcal{A} $ such that for any $A\in \mathcal{A} $ (possibly $A=A_{\text{final}}$, there is exactly one arrow from $A$ to $A_{\text{final}}$. 
\end{definition}
For example, the terminal object of $\mathsf{Set} $ is $\{*\} $, $\mathsf{Grp} $ is $\{1\} $, $\mathsf{CRing} $ is the zero ring, $\mathsf{Top} $ is the single point space, and a poset its maximal element (if one exists).

\subsection{Monomorphisms and epimorphisms}

\footnote{Here Evan uses the terminology ``\emph{monic}'' and ``\emph{epic}'', but I've noticed no one else really does that, so I'm replacing each instance with ``\emph{monomorphism}'' and ``\emph{epimorphism}''.}
Injectivity and surjectivity don't really make sense when talking about categories, because morphisms need not be functions. Here's the correct categorical notion:
\begin{definition}[Monomorphisms]
    A map $X \overset{f}{\to } Y$ is a \textbf{monomorphism} (or monic) if for any commutative diagram
\begin{figure}[H]
\centering
\begin{tikzcd}
A \arrow[r, "g", shift left] \arrow[r, "h"', shift right] & X \arrow[r, "f"] & Y
\end{tikzcd}
\end{figure}
we must have $g=h$. In other words, $f\circ g=f\circ h$ implies that $g=h$. \end{definition}
In a concrete category, injective implies monic: what the heck even is a concrete category? Anyway, consider $f\circ g=f\circ h$, so $f(g(a))=f(h(a)$ for all $a\in A$: but since $f$ is injective, this implies that $g(a)=h(a)$, and so $g=h$ and $f$ is a monomorphism. 
Similarly, the composition of two monomorphisms is also a monomorphism: let $f,g$ be monomorphisms. Then $(f\circ g)\circ \alpha = (f\circ g)\circ \alpha' \implies f \circ (g\circ \alpha )=f\circ (g\circ \alpha ') $ by associativity of arrows. Since $f$ is a monomorphism, $g\circ \alpha =g\circ \alpha '$, but since $g$ is also a monomorphism, $\alpha =\alpha '$ and we are done.
In most but not all situations, the converse of the definition also holds. For example, in $\mathsf{Set}, \mathsf{Grp} ,$ and $ \mathsf{CRing}  $, monic implies injective.

There are many categories with a ``free'' object that you can think of as elements. For example, an element of a set is a function $1 \to S$, and an element of a ring is a function $\Z[x]\to R$, etc. In all these categories, the definition of monomorphims literally say that ``$f$ is injective on $\operatorname{Hom}_{\mathcal{A} }(A,X)$''. However, there is a standard counterexample involving the category of ``divisible'' abelian groups $\mathsf{DivAbGrp} $ and the projection $\Q\to \Q /\Z$.
\begin{definition}[Epimorphisms]
    A map $X \overset{f}{\to } Y$ is an \textbf{epimorphism} (or epic) if for any commutative diagram 
    \begin{figure}[H]
    \centering
    \begin{tikzcd}
X \arrow[r, "f"] & Y \arrow[r, "g", shift left] \arrow[r, "h"', shift right] & A
\end{tikzcd}
    \end{figure}
    we must have $g=h$. In other words, $g\circ f=h\circ f\implies g=h$. \end{definition}
This is like surjectivity, but a little farther off. Furthermore, the correspondence failure rate is a little higher.
\begin{example}[Epimorphisms that aren't onto]
    \,
    \begin{enumerate}[label=(\alph*)]
        \item In $\mathsf{CRing} $, the inclusion $\Z\hookrightarrow \Q$ is an epimorphism that isn't onto. If two homomorphisms agree on an integer, they agree everywhere since we can extend linearly.
        \item In the category of \emph{Hausdorff} topological spaces, a mas is an epimorphism iff it has a dense image (for example $\Q\hookrightarrow \R$). 
    \end{enumerate}
    Basically, things fail when $f \colon X \to Y$ can be determined by just some of the points (some subset) of $X$.
\end{example}


\subsection{Functors}
Motivation: maps between categories, objects rising from other objects. 
\begin{example}[Basic Functors]
   Here are some basic examples of functors:
   \begin{itemize}
       \item Given an algebraic structure (group, field, vector space) we can take its underlying set $S$: this is a functor from $\mathsf{Grp}\to \mathsf{Set}$ (or whatever you want to start with).
       \item If we have a set $S$, if we consider the vector space with basis $S$ we get a functor $\mathsf{Set} \to \mathsf{Vect}$.
       \item Taking the power set of a set $S$ gives a functor $\mathsf{Set}\to \mathsf{Set}$.
       \item Given a locally small category $\mathcal{A}$, we can take a pair of objects $(A_1,A_2)$ and obtain a set $\operatorname{Hom}_{\mathcal{A}}(A_1,A_2)$. This turns out to be a functor $\mathcal{A}^{\text{op}}\times \mathcal{A}\to \mathsf{Set}$.
   \end{itemize}
   Finally, the most important example (WRT this course):
   \begin{itemize}
       \item In algebraic topology, we build groups like $H_1(X),\, \pi_1(X)$ associated to topological spaces. All these group constructions are functors $\mathsf{Top} \to \mathsf{Grp}$.
   \end{itemize}
\end{example}
\begin{definition}[Functors]
    Let $\mathcal{A}$ and $\mathcal{B}$ be categories. A \textbf{functor} $F$ takes every object of $\mathcal{A}$ to an object of $\mathcal{B}$. In addition, it must take every arrow $A_1 \overset{f}{\to }A_2$ to an arrow $F(A_1)\overset{F(f)}{\longrightarrow }F(A_2).$ Refer to the commutative diagram:
            \begin{figure}[H]
                \centering
                \begin{tikzcd}
               & A_1 \arrow[dd, "f"']       &  & B_1=F(A_1) \arrow[dd, "F(f)"] &                \\
\mathcal{A}\ni & {} \arrow[rr, "F", dotted] &  & {}                            & \in\mathcal{B} \\
               & A_2                        &  & B_2=F(A_2)                    &               
\end{tikzcd}
            \end{figure}
        Functors also satisfy the following requirements:
        \begin{itemize}
            \item Identity arrows get sent to identity arrows, that is, for each identity arrow $\operatorname{id}_A$, we have $F(\operatorname{id}_A)=\operatorname{id}_{F(A)}$.
            \item Functors respect composition: if $A_1\overset{f}{\to }A_2\overset{f}{\to }A_3$ are arrows in $\mathcal{A}$, then $F(g\circ f)=F(g)\circ F(f)$.
        \end{itemize}
\end{definition}
More precisely, these are covariant functors. A contravariant functor $F$ reverses the direction of arrows, so that $F$ sends $f \colon A_1 \to A_2$ to $F(f) \colon F(A_2) \to F(A_1)$, and satisfies $F(g\circ f)=F(f)\circ F(g)$ instead. A category $\mathcal{A}$ has an opposite category $\mathcal{A}^{\text{op}}$ with the same objects and with $\mathcal{A}^{\text{op}}(A_1,A_2)=\mathcal{A}(A_2,A_1)$. A contravariant functor $F \colon \mathcal{A} \to \mathcal{B}$ is just a covariant functor $\mathcal{A}^{\text{op}}\to \mathcal{B}$. 

\begin{example}
    We have already talked about \textbf{free} and \textbf{forgetful} functors in Example 1.3: the forgetful functors are functors from spaces to sets (the underlying set of a group) and free functors are from sets to spaces (the basis set forming a vector space).
    \begin{itemize}
        \item Another example of a forgetful functor is a functor $\mathsf{CRing}\to \mathsf{Grp}$ by sending a ring $R$ to its abelian group $(R,+)$.
        \item Another example of a free functor is a functor $\mathsf{Set}\to \mathsf{Grp}$ by taking the free group generated by a set $S$ (who would have known this is free?)
    \end{itemize}
\end{example}
Here is a cool example: functors preserve isomorphism. If two groups are isomorphic, then they must have the same cardinality. In the language of category theory, this can be expressed as such: if $G\cong H$ in $\mathsf{Grp}$ and $U \colon \mathsf{Grp} \to \mathsf{Set}$ is the forgetful functor, then $U(G)\cong U(H)$. We can generalize this to \emph{any} functor and category!
\begin{theorem}
    If $A_1\cong A_2$ are isomorphic objects in $\mathcal{A}$ and $F \colon \mathcal{A} \to \mathcal{B}$ is a functor then \[
        F(A_1)\cong F(A_2).
    \] 
\end{theorem}
\begin{proof}
Let's go diagram chasing!
            \begin{figure}[H]
                \centering
                \begin{tikzcd}
               & A_1 \arrow[dd, "f"', shift right=2]      &  & B_1=F(A_1) \arrow[dd, "F(f)"']                &                \\
\mathcal{A}\ni & {} \arrow[rr, "F", dotted, shift left=3] &  & {}                                            & \in\mathcal{B} \\
               & A_2 \arrow[uu, "g"']                     &  & B_2=F(A_2) \arrow[uu, "F(g)"', shift right=2] &               
\end{tikzcd}
            \end{figure}
The main idea of the proof follows from the fact that functors preserve composition and the identity map. 
\end{proof}
This is very very useful for us (people who are doing algebraic topology) because functors will preserve isomorphism between spaces (we get that homotopic spaces have isomorphic fundamental groups).
\begin{note}
    As a meme (or not really, but it's still funny), we can construct the category $\mathsf{Cat}$ whose objects are categories and arrows are functors.
\end{note}

\subsection{Homotopy Categories and Homotopy Equivalence}
Let $\mathsf{Top}_*$ be the category of pointed topological spaces. Then the fundamental group gives a functor $\mathsf{Top} _* \to \mathsf{Grp} $. When we have a suitable relation of homotopy between maps in a category $\mathcal{C} $, we define the homotopy category $\mathsf{Ho} (\mathcal{C} )$ to be the category sharing the same objects as $\mathcal{C} $, but morphisms the homotopy classes of maps. On $\mathsf{Top} _*$, we require homotopies to map basepoint to basepoint, and we get the homotopy category $\mathsf{hTop}_* $ of pointed spaces.

Homotopy equivalences in $\mathcal{C} $ are isomorphisms in $\mathsf{Ho} (\mathcal{C} )$. More concretely, recall that a map $f \colon X \to Y$ is a homotopy equivalence if there is a map $g \colon Y \to X$ such that both $g\circ f \simeq \operatorname{id}_X$ and $f\circ g \simeq \operatorname{id}_Y$. In the language of category theory, we can obtain the analogous notion of a pointed homotopy equivalence. Functors carry isomorphisms to isomorphisms, so then the pointed homotopy equivalence will induce an isomorphism of fundamental groups. This also holds, but less obviously, for the category of non pointed homotopy equivalences.
\begin{theorem}
    If $f \colon X \to Y$ is a homotopy equivalence, then \[
        f_* \colon \pi_1(X,x) \to \pi_1(Y,f(x))
    \] is an isomorphism for all $x\in X$.
\end{theorem}
\begin{proof}
    Let $g \colon Y \to X$ be a homotopy inverse of $f$. By our homotopy invariance diagram, we see that the composites \[
        \pi_1(X,x)\overset{f_*}{\longrightarrow }\pi_1(Y,f(x))\overset{g_*}{\longrightarrow}\pi_1(X,(g\circ f)(x))
    \] and \[
    \pi_1(Y,y)\overset{g_*}{\longrightarrow}\pi_1(X,g(y))\overset{f_*}{\longrightarrow}\pi_1(Y,(f\circ g)(y))
\] are isomorphisms determined by paths between basepoints given by chosen homotopies $g\circ f \simeq \operatorname{id}_X$ and $f\circ g \simeq \operatorname{id}_Y$. Then in each displayed composite, the first map is a monomorphism and the second is an epimorphism. Taking $y=f(x)$ in the second composite, we see that the second map in the first composite is an isomorphism. Therefore so is the first map, and we are done.
\end{proof}
A space $X$ is said to be contractible if it is homotopy equivalent to a point.
\begin{cor}
    The fundamental group of a contractible space is zero.
\end{cor}

\subsection{Natural Transformations}
We talked about maps between objects which led to categories, and then maps between categories which lead to functors. Now let's talk about maps between functors, the natural transformation: this is actually not too strange (recall the homotopy, a ``deformation'' from a map to another map).

In this case, we also want to pull a map (functor) $F$ to another map $G$ by composing a bunch of arrows in the target space $\mathcal{B}$.
\begin{definition}[Natural Transformations]
    Let $F,G \colon \mathcal{A} \to \mathcal{B}$ be two functors. A \textbf{natural transformation} $\alpha \colon F \to G$ denoted
            \begin{figure}[H]
                \centering
    \begin{tikzcd}
        \mathcal{A} \arrow[rr, "F", bend left] \arrow[rr, "G"', bend right] & \,\,~ \Big\Downarrow\alpha & \mathcal{B}
\end{tikzcd}
            \end{figure}
            consists of, for each $A\in \mathcal{A}$ an arrow $\alpha_A \in \operatorname{Hom}_{\mathcal{B}}(F(A),G(A))$, which is called the component of $\alpha$ at $A$. 
            Pictorially, it looks like this: 
            \begin{figure}[H]
                \centering
                \begin{tikzcd}
                                                                   &  & F(A)\in\mathcal{B} \arrow[dd, "\alpha_A"] \\
\mathcal{A}\ni A \arrow[rru, "F", dotted] \arrow[rrd, "G", dotted] &  &                                           \\
                                                                   &  & G(A)\in\mathcal{B}                       
\end{tikzcd}
            \end{figure}
            The $\alpha_A$ are subject to the ``naturality'' requirement such that for any $A_1 \overset{f}{\to }A_2$, the following diagram commutes:
            \begin{figure}[H]
                \centering
                \begin{tikzcd}
F(A_1) \arrow[r, "F(f)"] \arrow[d, "\alpha_{A_1}"'] & F(A_2) \arrow[d, "\alpha_{A_2}"] \\
G(A_1) \arrow[r, "G(f)"']                           & G(A_2)                          
\end{tikzcd}
            \end{figure}
            The arrow $\alpha_A$ represents the path that $F(A)$ takes to get to $G(A)$ (like in a homotopy from $f$ to $g$ the point $f(t)$ gets deformed to the point $g(t)$ continuously). Think of $f$ representing the homotopy and the basepoints being $F(A_1),G(A_1)$ to $F(A_2),G(A_2)$.
\end{definition}
Natural transformations can be composed. Take two natural transformations $\alpha \colon F \to G$ and $\beta \colon G \to H$. Consider the following commutative diagram: 
            \begin{figure}[H]
                \centering
                \begin{tikzcd}
                                                                                                     & F(A) \arrow[d, "\alpha_A"] \\
\mathcal{A}\ni A \arrow[ru, "F", dotted] \arrow[r, "G" description, dotted] \arrow[rd, "H"', dotted] & G(A) \arrow[d, "\beta_A"]  \\
                                                                                                     & H(A)                      
\end{tikzcd}
            \end{figure}
    We can also construct inverses: suppose $\alpha$ is a natural transformation such that $\alpha_A$ is an isomorphism for each $A$. Then we construct an inverse arrow $\beta_A$ in the following way:
            \begin{figure}[H]
                \centering
                \begin{tikzcd}
                                                                   &  & F(A)\in\mathcal{B} \arrow[dd, "\alpha_A"', shift right] \\
\mathcal{A}\ni A \arrow[rru, "F", dotted] \arrow[rrd, "G", dotted] &  &                                                         \\
                                                                   &  & G(A)\in\mathcal{B} \arrow[uu, "\beta_A"', shift right] 
\end{tikzcd}
            \end{figure}
            We say $\alpha$ is a \textbf{natural isomorphism}. Then $F(A)\cong G(A)$ \emph{naturally} in $A$ (and $\beta$ is an isomorphism too!) We write $F\cong G$ to show that the functors are naturally isomorphic.
\begin{example}
    If $F \colon \mathsf{Set} \to \mathsf{Grp}$ is the free functor that sends a set to the free group on such set and $U \colon \mathsf{Grp} \to \mathsf{Set}$ is the forgetful functor sending a free group to its generating set, then we have a natural inclusion of $S\hookrightarrow UF(S)$. The functors $F$ and $U$ are left and right adjoint to each other, in the sense that we have a natural isomorphism \[
        \mathsf{Grp}(F(S),A)\cong \mathsf{Set}(S,U(A))
    \] for a set $S$ and an abelian group $A$. This expresses the ``universal property'' of free objects: a map of sets $S\to U(A)$ extends uniquely to a homomorphism of groups $F(S)\to A$. 
\end{example}
\begin{definition}
    Two categories $\mathcal{A}$ and $\mathcal{B}$ are equivalent if there are functors $F \colon \mathcal{A} \to \mathcal{B}$ and $G \colon \mathcal{B} \to \mathcal{A}$ and natural isomorphisms $FG \to \operatorname{Id}$ and $GF \to \operatorname{Id}$, where the $\operatorname{Id}$ are the respective identity functors.
\end{definition}

\subsection{The Yoneda lemma (todo)}
\begin{definition}[The functor category]
    The \textbf{functor category} of two categories $\mathcal{A} $ and $\mathcal{B} $, denoted $[\mathcal{A} ,\mathcal{B} ]$ is defined as follows:
    \begin{itemize}
        \item The objects of $[\mathcal{A} ,\mathcal{B} ]$ are (covariant) functors $F \colon \mathcal{A}  \to \mathcal{B} $, and 
        \item The morphisms are natural transformations $\alpha \colon F \to G$.
    \end{itemize}
\end{definition}
todo

\subsection{Limits and colimits (todo)}
Let $\mathcal{D} $ be a small category and $\mathcal{C} $ be any category. A $\mathcal{D} $-shaped diagram in $\mathcal{C} $ is a functor $F \colon \mathcal{D}  \to \mathcal{C} $. A morphism $F\to F'$ of $\mathcal{D} $-shaped diagrams is a natural transformation, and we have the category $\mathcal{D} [\mathcal{C} ]$ of $\mathcal{D} $ shaped diagrams in $\mathcal{C} $. Any object $C$ of $\mathcal{C} $ determines the constant diagram
