\section{August 28, 2020: Homework 1}
\textbf{Section 1.1:} 8, 9, 25, \\
\textbf{Section 1.3:} 5, 15, \\
\textbf{Section 1.4:} 8, 10. \\
\\
\textbf{Section 2.1:} 5, 6, 12, 13, \\
\textbf{Section 2.3:} 2, 3, 12, 16, 20, \\
\textbf{Section 2.4:} 14, 15.\\
\\
\textbf{Section 3.1:} 3, 4, 14(a-c), 24, \\
\textbf{Section 3.2:} 8. 
\subsection{Problem 8 Section 1.1}
\begin{prob}
    Let $G=\{z\in \C \mid z^{n}=1\ \text{for some} \ n \in \Z^{+}\}.$
   \begin{enumerate}
       \item[(a)] Prove that $G$ is a group under multiplication (roots of unity).
       \item[(b)] Prove that $G$ is not a group under addition.
   \end{enumerate} 
\end{prob}

\subsection{Problem 9}
\begin{prob}
    Let $G=\{a+b\sqrt{2} \in \R \mid a,b\in \Q\} $.
    \begin{enumerate}
       \item[(a)] Prove that $G$ is a group under addition.
       \item[(b)] Prove that the nonzero elements of $G$ are a group under multiplication.
    \end{enumerate}
\end{prob}

\subsection{Problem 25}
\begin{prob}
    Prove that if $x^2=1$ for all $x\in G$ then $G$ is abelian.
\end{prob}

\subsection{Problem 5 Section 1.3}
\begin{prob}
    Find the order of (1 12 8 10 4)(2 13)(5 11 7)(6 9).
\end{prob}

\subsection{Problem 15}
\begin{prob}
    Prove that the order of an element in $S_n$ equals the least common multiple of the lengths of the cycles in its cycle decomposition. [Use Exercise 10 and Exercise 24 of Section 1.]
\end{prob}

\subsection{Problem 8 Section 1.4}
\begin{prob}
    Show that $\operatorname{GL}_n(F)$ is non-abelian for any $n \geq 2$ and any $F$.
\end{prob}

\subsection{Problem 10}
\begin{prob}
    Let $G=\{\left( \begin{smallmatrix}a&b\\ 0&c\end{smallmatrix}\right) \mid a,b,c,\in \R, \, a \neq 0, \, c \neq 0\}.$ 
    \begin{enumerate}
        \item[(a)] Compute the product of $\left( \begin{smallmatrix}a_1&b_1\\ 0&c_1\end{smallmatrix}\right)$ and $\left( \begin{smallmatrix}a_2&b_2\\ 0&c_2\end{smallmatrix}\right)$ to show that $G$ is closed under matrix multiplication.
        \item[(b)] Find the matrix inverse of $\left( \begin{smallmatrix}a&b\\ 0&c\end{smallmatrix}\right)$ and deduce that $G$ is closed under inverses.
        \item[(c)] Deduce that $G$ is a subgroup of $\operatorname{GL}_2(\R)$.
        \item[(d)] Prove that the set of elements of $G$ whos two diagonal entries are equal (i.e., $a=c$) is also a subgroup of $\operatorname{GL}_2(\R)$. 
    \end{enumerate}
\end{prob}

\subsection{Problem 5 Section 2.1}
\begin{prob}
    Prove that $G$ cannot have a subgroup $H$ with $|H|=n-1$, where $n=|G|>2$.
\end{prob}

\subsection{Problem 6}
\begin{prob}
    Let $G$ be an abelian group. Prove that $\{g\in G \mid |g| < \infty\} $ is a subgroup of $G$ (called the \emph{torsion subgroup} of $G$). Give an explicit example where this set is not a subgroup when $G$ is non-abelian.
\end{prob}

\subsection{Problem 12}
\begin{prob}
    Let $A$ be an abelian group and fix some $n\in \Z$. Prove that the following sets are subgroups of $A$: 
    \begin{enumerate}
        \item[(a)] $\{a^{n} \mid a\in A\} $,
        \item[(b)] $\{a\in A \mid a^{n}=1\} $.
    \end{enumerate}
\end{prob}

\subsection{Problem 13}
\begin{prob}
    Let $H$ be a subgroup of the additive group of rational numbers with the property that $1/x \in H$ for every nonzero element $x$ of $H$. Prove that $H=0$ or $\Q$.
\end{prob}

\subsection{Problem 2 Section 2.3}
\begin{prob}
    If $x$ is an element of the finite group $G$ and $|x|=|G|$, prove that $G=\langle x \rangle .$ Give an explicit example to show that this result need not be true if $G$ is an infinite group.
\end{prob}

\subsection{Problem 3}
\begin{prob}
    Find all generators for $\Z /48\Z$.
\end{prob}

\subsection{Problem 12}
\begin{prob}
    Prove that the following groups are \emph{not} cyclic.
    \begin{enumerate}
        \item[(a)] $\Z_2\times \Z_2$
        \item[(b)] $\Z_2\times \Z$
        \item[(c)] $\Z\times \Z$.
    \end{enumerate}
\end{prob}

\subsection{Problem 16}
\begin{prob}
    Assume $|X|=n$ and $|y|=m$. Supposed that $x$ and $y$ \emph{commute}: $xy=yx$. Prove that $|xy|$ divides the least common multiple of $m$ and $n$. Need this be true if $x$ and $y$ do \emph{not} commute? Give an example of commuting elements $x,y$ such that the order of $xy$ is not equal to the least common multiple of $|x|$ and $|y|$.
\end{prob}

\subsection{Problem 20}
\begin{prob}
    Let $p$ be a prime and $n$ be a positive integer. Show that if $x$ is an element of the group $G$ such that $x^{p^n}=1$ then $|x|=p^{m}$ for some $m\leq n$.
\end{prob}

\subsection{Problem 14 Section 2.4}
\begin{prob}
    A group $H$ is called \emph{finitely generated} is there is a finite set $A$ such that $H=\langle A \rangle $. 
    \begin{enumerate}
        \item[(a)] Prove that every finite group is finitely generated.
        \item[(b)] Prove that $\Z$ is finitely generated.
        \item[(c)] Prove that every finitely generated subgroup of the additive group $\Q$ is cyclic. [If $H$ is a finitely generated subgroup of $\Q$, show that $H \leq \langle \frac{1}{k} \rangle $, where $k$ is the product of all the denominators which appear in a set of generators for $H$.]
        \item[(d)] Prove that $\Q$ is not finitely generated.
    \end{enumerate}
\end{prob}

\subsection{Problem 15}
\begin{prob}
    Exhibit a proper subgroup of $\Q$ that is not cyclic.
\end{prob}

\subsection{Problem 3 Section 3.1}
\begin{prob}
    Let $A$ be an abelian group and let $B$ be a subgroup of $A$. Prove that $A/B$ is abelian. Give an example of a non-abelian group $G$ containing a proper normal subgroup $N$ such that $G/N$ is abelian.
\end{prob}

\subsection{Problem 4}
\begin{prob}
    Prove that in the quotient group $G/N, \, (gN)^{\alpha}=g^{\alpha}N$ for all $\alpha\in \Z$.
\end{prob}
 
\subsection{Problem 14}
Consider the additive quotient group $\Q/\Z$. 
\begin{enumerate}
    \item[(a)] Show that every coset of $\Z$ in $\Q$ contains exactly one representative $q\in \Q$ in the range $0\leq q < 1$.
    \item[(b)] Show that every element of $\Q/\Z$ has finite order but that there are elements of arbitrarily large order.
    \item[(c)] Show that $\Q/\Z$ is the torsion subgroup of $\R/\Z$.
\end{enumerate}

\subsection{Problem 24}
\begin{prob}
    Prove that $N \trianglelefteq G$ and $H$ is any subgroup of $G$ then $N \cap H \trianglelefteq H$.
\end{prob}

\subsection{Problem 8 Section 3.2}
\begin{prob}
    Prove that if $H$ and $K$ are finite subgroups of $G$ whose orders are relatively prime then $H \cap K = 1$. 
\end{prob}



