\section{First Order Linear Differential Equations (8/27/20)}
\subsection{Definitions}
\begin{definition}[Order]
    We have the \emph{order} of a differential equation the highest derivative of a function $y$ that appears in the equation. For example, the order of the differential equation \[
        \frac{dy}{dt}=3y^2\sin(t+y)
    \]
    is $1$, while the order of \[
    \frac{d^3y}{dt^3}=e^{-y}+t+\frac{d^2y}{dt^2}
    \]
    is $3$. We would call the first example a \emph{first-order differential equation} and the second a \emph{third order differential equation}.
\end{definition}
\begin{definition}[Solution]
    The \emph{solution} of a differential equation is a continuous function $y(t)$ that together with its derivatives satisfies the given relationship.
\end{definition}
\begin{example}
    The function \[
        y(t)=2\sin t - \frac{1}{3}\cos 2t
    \]
    is the solution of the second-order differential equation \[
        \frac{d^2y}{dt^2}+y= \cos 2t
    \]
    since 
    \begin{align*}
        \frac{d^2}{dt^2}\left( 2\sin t - \frac{1}{3}\cos 2t \right) + \left( 2\sin t - \frac{1}{3}\cos 2t \right) \\
        = \left( -2\sin t + \frac{4}{3} \cos 2t \right) + 2 \sin t - \frac{1}{3}\cos 2t = \cos 2t.
    \end{align*}
\end{example}

Goal: Given a differential equation of the form \[
    \frac{dy}{dt}=f(t,y)
\]
and the function $f(t,y)$, find all functions $y(t)$ that satisfy the equation above.

What we have: As of now, all we can solve is a differential equation of the form \[
    \frac{dy}{dt}=g(t)
\]
given $g(t)$ is integrable. Very sad!
\begin{definition}[Linear ODE]
    The general first-order linear differential equation is of the form \[
        \frac{dy}{dt}+a(t)y=b(t),
    \]
    where $a(t)$ and $b(t)$ are continuous (assumed to be functions of time).
\end{definition}

\subsection{Homogeneous Linear Ordinary Differential Equations}
\begin{definition}[Homogeneous Linear ODE]
    The equation \[
        \frac{dy}{dt}+a(t)y=0
    \]
    is called the \emph{homogeneous} first-order linear differential equation, and the previous definition is called the \emph{nonhomogeneous} first-order linear differential equation for $b(t)$ not necessarily zero.
\end{definition}
\begin{example}
    Let us solve the homogeneous first-order linear differential equation. Rewrite it in the form \[
        \frac{\frac{dy}{dt}}{y}=-a(t).
    \]
    Second, note that \[
        \frac{\frac{dy}{dt}}{y}\equiv \frac{d}{dt}\ln |y(t)|.
    \] Then we can write the differential equation in the form \[
    \frac{d}{dt}\ln |y(t)| = -a(t),
    \] so we have \[
    \ln|y(t)|= - \int a(t) \, dt + c_1.
    \] Continuing on, \[
    |y(t)|=\exp\left( -\int a(t) \, dt + c_1 \right) = c \exp \left( - \int a(t) \, dt \right) 
    \]
    or \[
        \left| \, y(t)\exp\left( \int a(t) \, dt \right)  \right| = c.
    \]
    Now $y(t)\exp \left( \int a(t) \, dt \right) $ is continuous and we know its absolute value is constant which implies that the function itself is constant (which follows from the IVT, assuming $g(t_1)=c$ and $g(t_2)=-c$ for $g$ a function, $c$ a constant. So we have $y(t)\exp\left( \int a(t) \, dt \right) = c$, or 
    \begin{equation}\label{eq 1}
        y(t) = c \exp \left( - \int a(t) \, dt \right) .
    \end{equation}
    Equation \eqref{eq 1} is the \emph{general solution} of the homogeneous equation. Note that there exist infinitely many solutions since for all $c$ we have a distinct $y(t)$.
\end{example}
\begin{example}
    To solve the Linear ODE \[
    \frac{dy}{dt}+2ty=0,
    \]
    simply apply Equation \eqref{eq 1} to yield \[
        y(t)=c\exp\left( -\int 2t \, dt \right) = c\exp\left(-t^2 \right) .
    \] (This is taking too long! I'll type notes with less rigor next time).
\end{example}

\subsection{Initial Value Problems}
Usually scientists are not interested in the general solution given by Equation \eqref{eq 1}, rather we look for solutions to a specific $y(t)$ which at some time $t_0$ has the value $y_0$, or we want to determine a $y(t)$ such that \[
    \frac{dy}{dt}+a(t)y=0, \quad y(t_0)=y_0.
\]
Please accept the derivation that a general solution to this type of problem is 
\begin{equation}\label{eq_2}
    y(t)=y_0\exp\left( -\int_{t_0}^{t} a(s) \, ds \right)  
\end{equation}
without proof (there is nothing of interest about the derivation process).
\begin{example}
    To solve \[
        \frac{dy}{dt}+(\sin t)y=0, \quad y(0)=\frac{3}{2},
    \]
    let  $a(t)=\sin t$, $t_0=0$, $y_0=\frac{3}{2}$. Then \[
        y(t)=\frac{3}{2}\exp\left( -\int_{0}^{t} \sin s \, ds \right)  =\frac{3}{2} \exp \left( \cos t - 1 \right).
    \]
\end{example}
\begin{example}
    To solve the initial value problem \[
        \frac{dy}{dt}+ \exp(t^2)y=0, \quad y(1)=2,
    \]
    simply PLUG IT IN (reee) to get \[
        y(t)=2\exp\left( -\int_{1}^{t} e^{s^2} \, ds \right).
    \]
    Recall that this is the \emph{Gaussian Integral} and can be solved by a change to double integration by polar coordinates (yielding $\int_{-\infty}^{\infty} e^{-x^2} \, dx = \sqrt{\pi}$), but in general has no closed form solution.
\end{example}

\subsection{Non-homogeneous Linear Differential Equations}
Recall the non-homogeneous linear differential equation of the form $\frac{dy}{dt}+a(t)y=b(t)$. Let $\mu(t)$ be a continuous function. Then multiply both sides by a continuous function $\mu(t)$ to get \[
    \mu(t)\frac{dy}{dt}+a(t)\mu(t)y=\mu(t)b(t),
\]
which is equivalent to the above form of a non-homogeneous linear differential equation. What we want: $\mu(t)\frac{dy}{dt}+a(t)\mu(t)y$ equal to the derivative of some simple expression. To get this, notice that \[
    \frac{d}{dt} \mu(t)y=\mu(t)\frac{dy}{dt}+\frac{d\mu}{dt}y
\]
by the product rule, so if $\frac{d\mu(t)}{dt}=a(t)\mu(t)$, our expression above will simply be equal to the derivative of $\mu(t)y$. Since our new expression is just a linear homogeneous differential equation, we have \[
    \mu(t)=\exp\left( \int a(t) \, dt \right) .
\]
Now the expressions \[
    \frac{d}{dt} \mu(t)y=\mu(t)\frac{dy}{dt}+\frac{d\mu}{dt}y \\
\] and \[
\frac{d}{dt} \mu(t)y=\mu(t)b(t)
\]
are equivalent, so we can integrate both sides to obtain \[
    \mu(t)y=\int \mu(t)b(t) \, dt+c
\]
or \[
    y=\frac{1}{\mu(t)}\left( \int \mu(t)b(t) \, dt +c\right)=\exp\left( -\int a(t) \, dt \right) \left( \int \mu(t)b(t) \, dt+c \right) .
\]
A similar integration between $t_0$ and $t$ yields \[
    \mu(t)y-\mu(t_0)y_0=\int_{t_0}^{t} \mu(s)b(s) \, ds
\] or \[
y=\frac{1}{\mu(t)}\left( \mu(t_0)y_0+\int_{t_0}^{t} \mu(s)b(s) \, ds \right),
\]
solving initial-value problems. 
\subsection{Integrating Factor}
Remark: $\mu(t)$ is called an \emph{integrating factor} for the nonhomogeneous equation since after multiplying both sides by $\mu(t)$ we can immediately integrate to find all solutions.
\begin{example}
    We find the general solution of the differential equation \[
    \frac{dy}{dt}-2ty=t.
    \]
    We know the integrating factor $\mu(t)$ is equal to $\exp\left( \int -2t \, dt \right) = e^{-t^2},$ so multiplying both sides by $\mu(t)$ yields \[
        e^{-t^2}\left( \frac{dy}{dt}-2ty \right) =e^{-t^2}t
    \] which is equivalent to \[
    \frac{d}{dt} \left( e^{-t^2}y \right) = e^{-t^2}t
    \]
    by our choice of $\mu(t)$. So \[
        \int \frac{d}{dt} \left( e^{-t^2}y \right)  \, dt = \int t e^{-t^2} \, dt, 
    \]
    and by the Fundamental Theorem we have \[
    e^{-t^2}y=-\frac{1}{2} e^{-t^2} + c.
    \]
     Finally, we conclude that \[
     y=-\frac{1}{2}+ce^{t^2}.
     \]
\end{example}
\begin{example}
    Here we solve an initial value problem. Let \[
        \frac{dy}{dx}+xy=xe^{x^2/2}, \quad y(0)=1.
    \] The integrating factor $\mu(x)$ is equal to $\exp\left( \int x\, dx \right) = e^{x^2/2}$. Multiply both sides by $\mu(x)$ to obtain \[
    e^{x^2/2}\left( \frac{dy}{dx}+xy \right) =e^{x^2}\left( xe^{x^2/2} \right),
    \] which is equivalent to
    \[
        \frac{d}{dx}\left( e^{x^2/2} y \right) = xe^{x^2}
    \]
    by our choice of $\mu(x)$. By the Fundamental Theorem, we have \[
    e^{x^2/2}y=\int xe^{x^2} \, dx=\frac{1}{2}e^{x^2}+c.
    \]
    Let $x=0$ and $y=1$, then  $1=\frac{1}{2}+c$, so $c=\frac{1}{2}$. We conclude that \[
        y(x)=\frac{1}{2}e^{x^2/2}+\frac{1}{2}e^{-x^2/2}.
\] We can simplify this to \[
y(x) = \frac{1}{2}e^{x^2/2}\left( 1+e^{-x^2} \right) .
\]
\end{example}

