\documentclass[fontsize=9pt]{article}
\input{../preamble.tex}
\title{De Rham Cohomology and Characteristic Classes Notes}
\usepackage{epigraph}
\begin{document}
\maketitle
I'm excited to say that I'm participating in the Directed Reading Program (DRP) this semester, mentored by Arun Debray! (Read more here: \url{web.ma.utexas.edu/users/drp}). This semester, I'm following a book called \emph{From Calculus to Cohomology: De Rham cohomology and characteristics classes} by Madsen and Tornehave. 

These are the full version of the notes, taken to help me learn the material. I plan on summarizing my results in a sort of exposition style to put on the DRP website, as well as a beamer presentation for the symposium. I plan on having all three files hosted on my website somewhere, probably around here: \url{https://git.simonxiang.xyz/math_notes/files.html}

\begin{center}
    \large\sc Prerequisites
\end{center}
The reader should be familiar with multivariable calculus and linear algebra at the minimum, as well as basic group theory (up to the first isomorphism theorem). Some things that are helpful but not necessary include:
\begin{itemize}
    \setlength\itemsep{-0.2em}
    \item Basic analysis, including open and closed sets, and the inverse function theorem.
    \item Point-set topology would be very nice.
    \item Algebraic topology would be very helpful, but I assume no knowledge of cohomology. 
\end{itemize}
In general, these notes will be taken like you know what open sets are, properties of connected spaces, what a commutative diagram is, stuff like that (because they were taken to help me learn the material). But for the condensed paper, I plan on introducing everything I need (besides stuff from calculus and linear algebra), so they can be somewhat self contained.
\tableofcontents
    %include notes files
    \include{prelim}
    \include{cohomology}
    %add and swap as necessary
\end{document}
